---
title: "Linear Regression Assignment"
author: "Mendoza College of Business"
date: "02/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up

```{r}
library(ggplot2)
load("nba_data.rda") # Load data
```


The dataset we are using contains statistics on NBA games from 2010 to 2021, the features in the data are:

The statistics calculated are:

* fgm - Field goals made
* fga - Field goals attempted
* fg3m - 3 point shots made
* fg3a - 3 point shots attempted
* fg2m - 2 point shots made
* fg2a - 2 point shots attempted
* ftm - Free throws made
* fta - Free throws attempted
* oreb - Offensive rebounds
* dreb - Defensive rebounds
* treb - Total rebounds
* ast - Assists
* stl - Steals
* blk - Blocks
* tov - Turnovers
* pf - Personal fouls
* pts - Points scored
* pctfg - Field goal percentage
* pctfg2 - 2 point field goal percentage
* pctfg3 - 3 point field goal percentage
* pctft - Free throw percentage

These are calculated for the different aspects of the matchup denoted by the following adjustments to the variable names:

* _1_off - Offensive stats for team 1 in previous games
* _1_def - Defensive stats for team 1 in previous games (That is stats from team 1's opponents)
* _2_off - Offensive stats for team 2 in previous games
* _2_def - Defensive stats for team 2 in previous games (That is stats from team 2's opponents)

There is also an ELO rating function which can be used to provide a ranking of the teams:

https://medium.com/purple-theory/what-is-elo-rating-c4eb7a9061e0

The objective of our analysis is to determine the key factors involved in winning NBA games, so we will use point differential as our response variable.

```{r}
# Create response variable
nba_data$point_diff <- nba_data$team_1_points - nba_data$team_2_points
```

Before running the model we want to drop all of the unique identifiers for the games and the points used to calculate the differential:

```{r}
# Drop columns categorical
nba_model <- nba_data[,c(8:91, 94:96)]
```


# Assignment

Please complete the following tasks and answer the following questions:


* 1 - Create two visualizations which may reveal interesting relationships in the dataset. Describe what you see in the visualization. (2 Points) 

```{r}
g_1 <- ggplot(nba_model, # Set dataset 
              aes(y = point_diff, # Set y-axis as insurance charges 
                  x = fga_1_off)) + # Set x-axis as age.
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(y = "Point Differential", # Set plot labels
       x = "FGA 1 Game Before",
       title = "Point Diff v FGA 1 Game Before")

g_1 # Generate plot
```
There is seemingly no relationship between point differential and the field goals attempted 1 game before.

```{r}
g_2 <- ggplot(nba_model, # Set dataset 
              aes(y = point_diff, # Set y-axis as insurance charges 
                  x = pts_1_def)) + # Set x-axis as age.
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(y = "Point Differential", # Set plot labels
       x = "Pts Against 1 Game Before",
       title = "Point Diff v Pts Against 1 Game Before")

g_2 # Generate plot
```
There is a very weak negative relationship between points against in the game before and point diff.

* 2 - Run a linear regression model on the dataset using `point_diff` as the response variable and all other variables in the model data as explanatory variables. (1 Point)

```{r}
fit1 <- lm(point_diff ~ ., data = nba_model)


```



* 3 - What features in the model were significant (At 0.1% level)? Do these variables have a positive or negative effect on a teams change of winning a game? (1 Point)

```{r}

summary(fit1)
p_values <- summary(fit1)$coefficients[, "Pr(>|t|)"]
significant_terms <- summary(fit1)$coefficients[p_values < 0.001, ]
significant_terms


```

dreb_1_off, stl_1_off, ast_2_def and elo_1 have a postiive effect.
tov_1_off, ast_1_de and elo_2 have a negative effect.


* 4 - Visualize two potential interaction terms and their relationship with the response variable. Are these interaction terms likely to have a significant relationship with the response variable? (2 Points)

```{r}
# Create plot
g_2 <- ggplot(nba_model, # Set dataset 
              aes(y = point_diff,
                  color = cut(ast_1_off,3), # Set y-axis 
                  x = oreb_1_off)) + # Set x-axis 
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(y = "Points diff", # Set plot labels
       x = "Off Reb 1 game ago",
       title = "Rebs and Assists",
       color = "Off Assists 1 game ago")

g_2 # Generate plot
```
When there is a high number of offensive assists 1 game ago, the offensive rebounds from 1 game ago will have a mildly positive effect on point diff. When there are a mid-low number of offensive assists 1 game ago the offensive rebounds from 1 game ago have a mildly negative effect. The interaction between assists 1 game ago and offensive rebs 1 game ago can change the relationship with the points diff.
```{r}
# Create plot
g_1 <- ggplot(nba_model, # Set dataset 
              aes(x = tov_1_off,
                  color = cut(tov_2_off,3), # Set y-axis 
                  y = point_diff)) + # Set x-axis 
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
        
  labs(x = "turnovers off 1 game before", # Set plot labels
       y = "point diff",
       title = "Offensive Turnovers",
       color = "turnovers off 2 games before")

g_1 # Generate plot
```
When there is a high number of offensive turnovers 2 games ago, the offensive turnovers from 1 game ago will have a mildly positive effect on point diff. When there are a mid-low number of offensive turnovers 2 games ago the offensive trunovers from 1 game ago have a mildly negative effect. The interaction between turnovers 1 game vs 2 games ago can change the relationship with the points diff.



* 5 - Fit a linear regression model with the two interaction terms included. (1 Point)

```{r}
fit2 <- lm(point_diff ~ . + tov_1_off * tov_2_off * ast_1_off * oreb_1_off, nba_model)

summary(fit2)
p_values <- summary(fit2)$coefficients[, "Pr(>|t|)"]
significant_terms2 <- summary(fit2)$coefficients[p_values < 0.001, ]
significant_terms2
```


* 6 - How has the inclusion of the interaction terms affected the coefficients in the model? How has the inclusion of interaction terms affected the model fit? (1 Point)

The inclusion of interaction terms increased the R^2 value very slightly by 0.002 but none of the interaction terms were significant at a 0.1% significant level. It also decreased the number of other vars that were significant (tov_1_off is no longer significant) but did not change the direction of their coefficients.


2 Points for code quality and analysis decisions.
